---
title: "Manhattan Rolling Data Sales Paper"
author: "Kevin Cannon, Vishi Cline, Abhishek Dharwadkar"
date: "October 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description:  Markdown analysis on the cleaned datafile for manhattan rolling sales data

Load libraries gdata and plyr
```{r}
library(gdata)
library(plyr)
```

Read the cleaned comma delimited csv file whith header starting at row 1
```{r}
manhattan.clean <- read.csv("D:\\My Documents\\GitHub\\Manhattan-Rolling-Data-Sales\\Data\\2011_manhattan_tidy.csv",header=TRUE)
```

Check what variables are there and the their types
```{r}
str(manhattan.clean)
```

Explore the clean data with histograms
```{r}
hist(manhattan.clean$gross.sqft)
hist(log10(manhattan.clean$gross.sqft))
hist(manhattan.clean$sale.price.n)
hist(log10(manhattan.clean$sale.price.n))
```

Explore the data with scatterplots to check the effect of the size of the house on the sale price
```{r}
plot(log10(manhattan.clean$gross.sqft),log10(manhattan.clean$sale.price.n), pch = 16, cex = 1.3, col = "blue", xlab = "Gross Sqft", ylab = "Sale Price")
```

We can also explore the data by looking at the boxplots of the sale price by building class category
```{r}
boxplot(log10(manhattan.clean$sale.price.n)~manhattan.clean$building.class.category)
```